Fourier-domain modulations and delays of gravitational-wave signals

### 2018-07-23 docs/paper_method.tex

\begin{figure}
  \centering
  \includegraphics[width=.98\linewidth]{plots/tf_py.pdf}
  \caption{Time-to-frequency correspondence $t_{f}$, as defined directly from the phase of the Fourier-domain signal in~\eqref{eq:deftf}, in geometric units. We show only the high-frequency part of the signal, corresponding to the merger region. The PhenomD waveforms have been aligned such that the time-domain amplitude (obtained by an IFFT) peaks at $t=0$. The two aligned spins are equal, with components of $0.95$ ($++$, blue), $0$ ($00$, red), and $-0.95$ ($--$, yellow), for mass ratios $q=1$ (full line) and $q=8$ (dashed). The vertical lines shows the time-domain instantaneous frequency at the peak $\omega^{22}_{\rm peak}/(2\pi)$, and the fact that the curves $t_{f}$ do not pass exactly by the crossing with the $t=0$ horizontal line reflects the fact that the link between time-domain frequency and Fourier frequency $f$ is only approximate at merger. Note that $t_{f}$ increases to at most $\sim 30M$ after the peak, and is not monotonous.}
  \label{fig:tf}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[width=.98\linewidth]{plots/TfTA_py.pdf}
  \caption{Fourier-domain amplitude and phase timescales, as defined in~\eqref{eq:defTf} and~\eqref{eq:defTA}, in geometric units for an equal-mass, non-spinning system. The hierarchy of these timescales is roughly the same for higher mass ratio and higher spin systems. The time-domain frequency at merger and the ringdown frequency are represented by the thick and thin vertical lines. The derivatives from which these timescales are built encounter zero-crossings and change sign in the high-frequency range.}
  \label{fig:TfTA}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[width=.99\linewidth]{plots/lisafom_Psi_McDeltat_py.pdf}
  \caption{Contour levels for the analytical estimate of the error measure $\epsilon_{\Psi 2}$ at the starting frequency, as a function of chirp mass $\Mchirp$ and time to coalescence $\Delta t$. Blue corresponds to the orbital response and red to the constellation response. The colored shaded areas indicate regions where $\epsilon_{\Psi 2} \geq 1$, where the perturbative formalism is expected to break down. In the region to the right of the black line, $f_{\rm start}$ given in~\eqref{eq:fstartN} is lower than the lowest in-band frequency $f_{\rm min} = 10^{-5}\Hz$, so that the signal starts at $f_{\rm min}$ and $\epsilon_{\Psi 2}$ becomes independent of $\Delta t$.}
  \label{fig:lisafomPsiMcDeltat}
\end{figure}

\begin{figure*}
  \centering
  \includegraphics[width=.98\linewidth]{plots/lisafom_py.pdf}
  \caption{Error estimates as defined in~\eqref{eq:deffom}, for an equal-mass, non-spinning system and for total masses $M=10^{7} \Msol$, $10^{4} \Msol$ and $10^{2} \Msol$. The top row corresponds to the orbital delay part of the response~\eqref{eq:defresponse0}, and the lower row shows the LISA-centered constellation response~\eqref{eq:defresponseL}. The error measures $\epsilon_{\Psi 2}$ for the phase corrections, $\epsilon_{A1}$ for the amplitude, $\epsilon_{d}$ for the delay are shown from left to right. The central line and interval are the mean and $1\sigma$ standard deviation of the logarithm of $\epsilon$ computed with numerical derivatives over 400 random values for the position in the sky, inclination and polarization. The starting frequency is set by an observation time of $\Delta t = 10 \text{yrs}$ before merger. We overlay in dashed the analytical estimates obtained from from~\eqref{eq:timescalesNfstart} and~\eqref{eq:estimatederivorb}-\eqref{eq:estimatederivconst}.}
  \label{fig:fomLISA}
\end{figure*}

@@

\begin{figure}
  \centering
  \includegraphics[width=.98\linewidth]{plots/LISAtransfererrorM1e7dO_py.pdf}
  \caption{Transfer function and residual for the orbital response~\eqref{eq:defresponse0}, for an equal-mass system with $M=10^{7} \Msol$. The first and second panels show the amplitude and phase of the transfer function. The bottom panel shows the relative modulus of the residual of the perturbative treatment~\eqref{eq:summaryNAd}, with the color indicating the order of the approximation, compared to a numerical FFT. The starting frequency is set by the LISA sensitivity band at $10^{-5}\mathrm{Hz}$, corresponding to $3.8\mathrm{yr}$. The thick and thin vertical lines indicate the merger and ringdown frequencies.}
  \label{fig:LISAerrorM1e7orb}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=.98\linewidth]{plots/LISAtransfererrorM1e2dO_py.pdf}
  \caption{Transfer function and residual for the orbital response~\eqref{eq:defresponse0}, for an equal-mass system with $M=10^{2} \Msol$. The panels are as in Fig.~\ref{fig:LISAerrorM1e7orb}. The starting frequency corresponds to 10 years before merger. The ending frequency is set by the LISA sensitivity band at $1\mathrm{Hz}$, and the merger is out of band. The jump in the residuals for $\{N:3|A:1|d:1\}$ is due to the fact that we split the band in two, with two different sampling rates, for the FFT; it shows that, at this level, the details of the conditioning for the FFT start to cause numerical errors in the reference transfer function.}
  \label{fig:LISAerrorM1e2orb}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=.98\linewidth]{plots/LISAtransfererrorM1e7y12L_py.pdf}
  \caption{Transfer function and residual for the constellation response $y_{132}$~\eqref{eq:defresponseL}, for an equal-mass system with $M=10^{7} \Msol$. The top panel shows the real and imaginary parts of the transfer function, rescaled by the overall scaling $\pi f L$ of the response at low frequencies, while the bottom panel shows the relative residuals compared to an FFT. The frequencies are the same as in Fig.~\ref{fig:LISAerrorM1e7orb}.}
  \label{fig:LISAerrorM1e7const}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=.98\linewidth]{plots/LISAtransfererrorM1e2y12L_py.pdf}
  \caption{Transfer function and residual for the constellation response $y_{132}$~\eqref{eq:defresponseL}, for an equal-mass system with $M=10^{2} \Msol$. The top panel shows the real and imaginary parts of the transfer function, without the rescaling of Fig.~\ref{fig:LISAerrorM1e7const}. The bottom panels shows the residuals normalized by~\eqref{eq:transferLenvelope} to avoid zero-crossings. The frequencies are the same as in Fig.~\ref{fig:LISAerrorM1e2orb}, and the discontinuity in the residuals for $\{N:3|A:1|d:1\}$ is again due to the conditioning of the FFT.}
  \label{fig:LISAerrorM1e2const}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[width=.98\linewidth]{plots/LISACombExtent_py.pdf}
  \caption{Extent of the Fourier-domain comb entering the convolution~\eqref{eq:transferdiscreteconvolution}, determined from the criterion~\eqref{eq:criteriontruncationcomb} with truncation levels $\eta=10^{-12}$, $10^{-6}$, $10^{-3}$. Blue corresponds to the orbital delay comb coefficients $c_{n}^{0}$ given by~\eqref{eq:cn0}, and red to the constellation comb $c_{n}^{L}$ for the response~\eqref{eq:GslrL}. The result is shown after averaging over 100 random orientation angles. Most of the slowly-chirping systems considered here will have $f \lesssim 10^{-2}\Hz$.}
  \label{fig:lisacombextent}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=.98\linewidth]{plots/LISAtransfererrorSOBHepsPsi1d0corr_py.pdf}
  \caption{Transfer function and residual for the orbital delay $d_{0}$~\eqref{eq:delay0}, for an SOBH with $\epsilon_{\Psi 2}\sim 0.6$ for the orbital response. The frequency band corresponds to 4 years of observation for this system with $M=50\Msol$, $\Delta t = 200\yr$ away from merger. The top panels show the transfer function amplitude and phase at various orders of approximation in~\eqref{eq:summaryNAd}, with the numerical result of the FFT in black and the comb result~\eqref{eq:transferdiscreteconvolution} in green. The bottom panel shows the relative residuals with respect ot the FFT.}
  \label{fig:LISAerrorSOBHepsPsi1d0}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=.98\linewidth]{plots/LISAtransfererrorSOBHepsPsi10d0corr_py.pdf}
  \caption{Same as Fig.~\ref{fig:LISAerrorSOBHepsPsi1d0}, but for an SOBH with $\epsilon_{\Psi 2}\sim 4.5$ for the orbital response. The frequency band corresponds to 4 years of observation for this system with $M=15\Msol$, $\Delta t = 1500\yr$ away from merger. The perturbative treatment breaks down in this case, with errors of order 1.}
  \label{fig:LISAerrorSOBHepsPsi10d0}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=.98\linewidth]{plots/LISAtransfererrorSOBHepsPsi1y12L_py.pdf}
  \caption{Transfer function and residual for the constellation response $y_{132}$~\eqref{eq:GslrL}, for the same SOBH as in Fig.~\ref{fig:LISAerrorSOBHepsPsi1d0}. Here $\epsilon_{\Psi 2}\sim 0.01$ for the constellation part of the response. The top panel shows the real and imaginary part of the transfer function.}
  \label{fig:LISAerrorSOBHepsPsi1const}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=.98\linewidth]{plots/LISAtransfererrorSOBHepsPsi10y12L_py.pdf}
  \caption{Transfer function and residual for the constellation response $y_{132}$~\eqref{eq:GslrL}, for the same SOBH as in Fig.~\ref{fig:LISAerrorSOBHepsPsi10d0}. Here $\epsilon_{\Psi 2}\sim 0.1$ for the constellation part of the response. The top panel shows the real and imaginary part of the transfer function.}
  \label{fig:LISAerrorSOBHepsPsi10const}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[width=.98\linewidth]{plots/eulerSXS0058_py.pdf}
  \caption{Evolution of Euler angles $(\alpha, \gamma)$ (top panel) and $\beta$ (bottom panel) near merger for the example SpEC waveform SXS:BBH:0058~\cite{SpEC, SXScatalog, Mroue+12, Mroue+13}. Full, dashed and dotted lines represent the NR, SEOB and extended SEOB waveforms respectively. The vertical line at $t=0$ indicates the time of merger. In the left panel the full and dashed black lines indicate the asymptotic behaviour~\eqref{eq:OmegaframeQNM} for the rotation of the frame around the direction of the final $\bm{J}$. The difference between NR and SEOB there is due to the value of the final spin, with $\chi_{f}^{\rm NR} = 0.54$ and $\chi_{f}^{\rm SEOB} = 0.42$, which leads to a different $\Omega_{\rm frame}$. In the right panel, $\beta$ is asymptotically constant in our toy model.}
  \label{fig:precmodel}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[width=.98\linewidth]{plots/postmergeromega_py.pdf}
  \caption{Quasi-normal mode frequencies for the 0th overtone of the modes $22$ and $21$, as a function of the dimensionless spin of the final black hole $\chi_{f}$, as well as their difference $\Omega_{\rm frame}$ (see~\eqref{eq:OmegaframeQNM}). The lower panel shows the  ratio $\Omega_{\rm frame} / \omega^{\rm QNM}_{220}$ characteristic of the separation of timescales between the phase of the modulation~\eqref{eq:wignerphasesimpleprec} and the P-frame waveform phase.}
  \label{fig:QNM}
\end{figure}

@@

\begin{figure*}
  \centering
  \includegraphics[width=.98\linewidth]{plots/precfom_py.pdf}
  \caption{Error estimates of the approximation as defined in~\eqref{eq:deffom} for the precession modulations, for the three cases $++$, $\perp\perp$ and $--$ listed in Table~\ref{tab:precexamples}. The thick and thin vertical lines show, respectively, the merger frequency and the asymptotic ringdown frequency. The colors correspond to different values of $m$ in $\calT^{2}_{m 2}$ as defined in~\eqref{eq:defprectransfer}. The full lines are computed with numerical derivatives for the full waveform and modulation, while in the cases $++$ and $\perp\perp$ the dashed lines show the leading-order PN estimates~\eqref{eq:precfomPN} for single-spin simple precession. We show the three error estimates $\epsilon_{\Psi 2}$, $\epsilon_{A1}$, $\epsilon_{A2}$ defined in~\eqref{eq:deffom}, and the range $\epsilon \gtrsim 1$ indicates a breakdown of the formalism of Sec.~\ref{sec:formalism}.}
  \label{fig:fomprec}
\end{figure*}

@@

\begin{figure*}
  \centering
  \includegraphics[width=.98\linewidth]{plots/trigopoly_py.pdf}
  \caption{Amplitude (left panel) and phase (right panel), compared to its trigonometric polynomial representation for the three cases listed in Table~\ref{tab:precexamples}. The thick and thin vertical lines indicate the frequency of the merger and ringdown (the QNM frequency) respectively. The continuous line shows the target signal $\tilde{h}(f)$, while the dashed line shows the artificially symmetrized trigonometric polynomial~\eqref{eq:hsymtrigo}. The dots show the discrete samples entering~\eqref{eq:ykDFTprec}. The lower panels show the amplitude and phase residuals of~\eqref{eq:hsymtrigo} compared to the original $\tilde{h}(f)$.}
  \label{fig:trigopoly}
\end{figure*}

@@

\begin{figure*}
  \centering
  \includegraphics[width=.98\linewidth]{plots/precerror_++_py.pdf}
  \caption{Fourier-domain amplitudes of the modes $\tilde{h}^{\rm I}_{\ell m}$ (top panels), reconstruction errors in the transfer fonctions $\calT^{2}_{m2}$ for the amplitudes (middle panels) and for the phases (bottom panels),  for the configuration $++$ of Table~\ref{tab:precexamples}, with spins almost aligned. The thick and thin vertical lines show the merger and asymptotic ringdown frequencies respectively. For the solid curves in the top panel, transfer functions were computed from an FFT of the full time-domain precessing signal, while the dashed curves show the results of the Fourier-domain reconstruction techniques presented here. The columns show the orders of approximation $\{N:0 | A:0 | \text{No Conv.}\}$, $\{N:3 | A:2 | \text{No Conv.}\}$, $\{N:3 | A:2 | \text{Conv.}\}$. In the right column, the region of transition between the perturbative treatment of Sec.~\ref{subsec:executivesummary} and the high-frequency treatment of Sec.~\ref{subsec:trigopoly} is shaded in grey. Note that the errors shown here are relative to a given mode, not to the leading mode, and the errors are largest in this sense for the most subdominant modes; the top panels illustrate the mode hierarchy.}
  \label{fig:precerrors++}
\end{figure*}

\begin{figure*}
  \centering
  \includegraphics[width=.98\linewidth]{plots/precerror_pp_py.pdf}
  \caption{Same as Fig.~\ref{fig:precerrors++}, but for the configuration $\perp\perp$ of Table~\ref{tab:precexamples}, with both spins in the orbital plane.}
  \label{fig:precerrorspp}
\end{figure*}

\begin{figure*}
  \centering
  \includegraphics[width=.98\linewidth]{plots/precerror_--_py.pdf}
  \caption{Same as Fig.~\ref{fig:precerrors++}, but for the configuration $--$ of Table~\ref{tab:precexamples}, with almost anti-aligned spins.}
  \label{fig:precerrors--}
\end{figure*}

\begin{figure*}
  \centering
  \includegraphics[width=.98\linewidth]{plots/precMM_py.pdf}
  \caption{Unfaithfulness at various orders of approximation as compared with an FFT of the full time-domain precessing signal, for the three precessing systems of Table~\ref{tab:precexamples}. The color indicates the order of approximation of the perturbative treatment of Sec.~\ref{subsec:executivesummary}. Dashed lines are obtained applying the treatment of Sec.~\ref{subsec:executivesummary} to the whole frequency band, while full lines are obtained using Sec.~\ref{subsec:trigopoly} for the high frequency range. The computation is done using~\eqref{eq:defMM} with $f_{\rm min} = 20\mathrm{Hz}$, for masses between 20 and 400 $\Msol$, and with the aLIGO ZDHP noise curve~\cite{LIGOProspects13}. The inclination is defined here as the angle between the line of sight and the final $\bm{J}$.}
  \label{fig:precunfaithfulness}
\end{figure*}

>>>>

\begin{figure}
  \centering
  \includegraphics[width=.98\linewidth]{tf_py.pdf}
  \caption{Time-to-frequency correspondence $t_{f}$, as defined directly from the phase of the Fourier-domain signal in~\eqref{eq:deftf}, in geometric units. We show only the high-frequency part of the signal, corresponding to the merger region. The PhenomD waveforms have been aligned such that the time-domain amplitude (obtained by an IFFT) peaks at $t=0$. The two aligned spins are equal, with components of $0.95$ ($++$, blue), $0$ ($00$, red), and $-0.95$ ($--$, yellow), for mass ratios $q=1$ (full line) and $q=8$ (dashed). The vertical lines shows the time-domain instantaneous frequency at the peak $\omega^{22}_{\rm peak}/(2\pi)$, and the fact that the curves $t_{f}$ do not pass exactly by the crossing with the $t=0$ horizontal line reflects the fact that the link between time-domain frequency and Fourier frequency $f$ is only approximate at merger. Note that $t_{f}$ increases to at most $\sim 30M$ after the peak, and is not monotonous.}
  \label{fig:tf}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[width=.98\linewidth]{TfTA_py.pdf}
  \caption{Fourier-domain amplitude and phase timescales, as defined in~\eqref{eq:defTf} and~\eqref{eq:defTA}, in geometric units for an equal-mass, non-spinning system. The hierarchy of these timescales is roughly the same for higher mass ratio and higher spin systems. The time-domain frequency at merger and the ringdown frequency are represented by the thick and thin vertical lines. The derivatives from which these timescales are built encounter zero-crossings and change sign in the high-frequency range.}
  \label{fig:TfTA}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[width=.99\linewidth]{lisafom_Psi_McDeltat_py.pdf}
  \caption{Contour levels for the analytical estimate of the error measure $\epsilon_{\Psi 2}$ at the starting frequency, as a function of chirp mass $\Mchirp$ and time to coalescence $\Delta t$. Blue corresponds to the orbital response and red to the constellation response. The colored shaded areas indicate regions where $\epsilon_{\Psi 2} \geq 1$, where the perturbative formalism is expected to break down. In the region to the right of the black line, $f_{\rm start}$ given in~\eqref{eq:fstartN} is lower than the lowest in-band frequency $f_{\rm min} = 10^{-5}\Hz$, so that the signal starts at $f_{\rm min}$ and $\epsilon_{\Psi 2}$ becomes independent of $\Delta t$.}
  \label{fig:lisafomPsiMcDeltat}
\end{figure}

\begin{figure*}
  \centering
  \includegraphics[width=.98\linewidth]{lisafom_py.pdf}
  \caption{Error estimates as defined in~\eqref{eq:deffom}, for an equal-mass, non-spinning system and for total masses $M=10^{7} \Msol$, $10^{4} \Msol$ and $10^{2} \Msol$. The top row corresponds to the orbital delay part of the response~\eqref{eq:defresponse0}, and the lower row shows the LISA-centered constellation response~\eqref{eq:defresponseL}. The error measures $\epsilon_{\Psi 2}$ for the phase corrections, $\epsilon_{A1}$ for the amplitude, $\epsilon_{d}$ for the delay are shown from left to right. The central line and interval are the mean and $1\sigma$ standard deviation of the logarithm of $\epsilon$ computed with numerical derivatives over 400 random values for the position in the sky, inclination and polarization. The starting frequency is set by an observation time of $\Delta t = 10 \text{yrs}$ before merger. We overlay in dashed the analytical estimates obtained from from~\eqref{eq:timescalesNfstart} and~\eqref{eq:estimatederivorb}-\eqref{eq:estimatederivconst}.}
  \label{fig:fomLISA}
\end{figure*}

@@

\begin{figure}
  \centering
  \includegraphics[width=.98\linewidth]{LISAtransfererrorM1e7dO_py.pdf}
  \caption{Transfer function and residual for the orbital response~\eqref{eq:defresponse0}, for an equal-mass system with $M=10^{7} \Msol$. The first and second panels show the amplitude and phase of the transfer function. The bottom panel shows the relative modulus of the residual of the perturbative treatment~\eqref{eq:summaryNAd}, with the color indicating the order of the approximation, compared to a numerical FFT. The starting frequency is set by the LISA sensitivity band at $10^{-5}\mathrm{Hz}$, corresponding to $3.8\mathrm{yr}$. The thick and thin vertical lines indicate the merger and ringdown frequencies.}
  \label{fig:LISAerrorM1e7orb}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=.98\linewidth]{LISAtransfererrorM1e2dO_py.pdf}
  \caption{Transfer function and residual for the orbital response~\eqref{eq:defresponse0}, for an equal-mass system with $M=10^{2} \Msol$. The panels are as in Fig.~\ref{fig:LISAerrorM1e7orb}. The starting frequency corresponds to 10 years before merger. The ending frequency is set by the LISA sensitivity band at $1\mathrm{Hz}$, and the merger is out of band. The jump in the residuals for $\{N:3|A:1|d:1\}$ is due to the fact that we split the band in two, with two different sampling rates, for the FFT; it shows that, at this level, the details of the conditioning for the FFT start to cause numerical errors in the reference transfer function.}
  \label{fig:LISAerrorM1e2orb}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=.98\linewidth]{LISAtransfererrorM1e7y12L_py.pdf}
  \caption{Transfer function and residual for the constellation response $y_{132}$~\eqref{eq:defresponseL}, for an equal-mass system with $M=10^{7} \Msol$. The top panel shows the real and imaginary parts of the transfer function, rescaled by the overall scaling $\pi f L$ of the response at low frequencies, while the bottom panel shows the relative residuals compared to an FFT. The frequencies are the same as in Fig.~\ref{fig:LISAerrorM1e7orb}.}
  \label{fig:LISAerrorM1e7const}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=.98\linewidth]{LISAtransfererrorM1e2y12L_py.pdf}
  \caption{Transfer function and residual for the constellation response $y_{132}$~\eqref{eq:defresponseL}, for an equal-mass system with $M=10^{2} \Msol$. The top panel shows the real and imaginary parts of the transfer function, without the rescaling of Fig.~\ref{fig:LISAerrorM1e7const}. The bottom panels shows the residuals normalized by~\eqref{eq:transferLenvelope} to avoid zero-crossings. The frequencies are the same as in Fig.~\ref{fig:LISAerrorM1e2orb}, and the discontinuity in the residuals for $\{N:3|A:1|d:1\}$ is again due to the conditioning of the FFT.}
  \label{fig:LISAerrorM1e2const}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[width=.98\linewidth]{LISACombExtent_py.pdf}
  \caption{Extent of the Fourier-domain comb entering the convolution~\eqref{eq:transferdiscreteconvolution}, determined from the criterion~\eqref{eq:criteriontruncationcomb} with truncation levels $\eta=10^{-12}$, $10^{-6}$, $10^{-3}$. Blue corresponds to the orbital delay comb coefficients $c_{n}^{0}$ given by~\eqref{eq:cn0}, and red to the constellation comb $c_{n}^{L}$ for the response~\eqref{eq:GslrL}. The result is shown after averaging over 100 random orientation angles. Most of the slowly-chirping systems considered here will have $f \lesssim 10^{-2}\Hz$.}
  \label{fig:lisacombextent}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=.98\linewidth]{LISAtransfererrorSOBHepsPsi1d0corr_py.pdf}
  \caption{Transfer function and residual for the orbital delay $d_{0}$~\eqref{eq:delay0}, for an SOBH with $\epsilon_{\Psi 2}\sim 0.6$ for the orbital response. The frequency band corresponds to 4 years of observation for this system with $M=50\Msol$, $\Delta t = 200\yr$ away from merger. The top panels show the transfer function amplitude and phase at various orders of approximation in~\eqref{eq:summaryNAd}, with the numerical result of the FFT in black and the comb result~\eqref{eq:transferdiscreteconvolution} in green. The bottom panel shows the relative residuals with respect ot the FFT.}
  \label{fig:LISAerrorSOBHepsPsi1d0}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=.98\linewidth]{LISAtransfererrorSOBHepsPsi10d0corr_py.pdf}
  \caption{Same as Fig.~\ref{fig:LISAerrorSOBHepsPsi1d0}, but for an SOBH with $\epsilon_{\Psi 2}\sim 4.5$ for the orbital response. The frequency band corresponds to 4 years of observation for this system with $M=15\Msol$, $\Delta t = 1500\yr$ away from merger. The perturbative treatment breaks down in this case, with errors of order 1.}
  \label{fig:LISAerrorSOBHepsPsi10d0}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=.98\linewidth]{LISAtransfererrorSOBHepsPsi1y12L_py.pdf}
  \caption{Transfer function and residual for the constellation response $y_{132}$~\eqref{eq:GslrL}, for the same SOBH as in Fig.~\ref{fig:LISAerrorSOBHepsPsi1d0}. Here $\epsilon_{\Psi 2}\sim 0.01$ for the constellation part of the response. The top panel shows the real and imaginary part of the transfer function.}
  \label{fig:LISAerrorSOBHepsPsi1const}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=.98\linewidth]{LISAtransfererrorSOBHepsPsi10y12L_py.pdf}
  \caption{Transfer function and residual for the constellation response $y_{132}$~\eqref{eq:GslrL}, for the same SOBH as in Fig.~\ref{fig:LISAerrorSOBHepsPsi10d0}. Here $\epsilon_{\Psi 2}\sim 0.1$ for the constellation part of the response. The top panel shows the real and imaginary part of the transfer function.}
  \label{fig:LISAerrorSOBHepsPsi10const}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[width=.98\linewidth]{eulerSXS0058_py.pdf}
  \caption{Evolution of Euler angles $(\alpha, \gamma)$ (top panel) and $\beta$ (bottom panel) near merger for the example SpEC waveform SXS:BBH:0058~\cite{SpEC, SXScatalog, Mroue+12, Mroue+13}. Full, dashed and dotted lines represent the NR, SEOB and extended SEOB waveforms respectively. The vertical line at $t=0$ indicates the time of merger. In the left panel the full and dashed black lines indicate the asymptotic behaviour~\eqref{eq:OmegaframeQNM} for the rotation of the frame around the direction of the final $\bm{J}$. The difference between NR and SEOB there is due to the value of the final spin, with $\chi_{f}^{\rm NR} = 0.54$ and $\chi_{f}^{\rm SEOB} = 0.42$, which leads to a different $\Omega_{\rm frame}$. In the right panel, $\beta$ is asymptotically constant in our toy model.}
  \label{fig:precmodel}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[width=.98\linewidth]{postmergeromega_py.pdf}
  \caption{Quasi-normal mode frequencies for the 0th overtone of the modes $22$ and $21$, as a function of the dimensionless spin of the final black hole $\chi_{f}$, as well as their difference $\Omega_{\rm frame}$ (see~\eqref{eq:OmegaframeQNM}). The lower panel shows the  ratio $\Omega_{\rm frame} / \omega^{\rm QNM}_{220}$ characteristic of the separation of timescales between the phase of the modulation~\eqref{eq:wignerphasesimpleprec} and the P-frame waveform phase.}
  \label{fig:QNM}
\end{figure}

@@

\begin{figure*}
  \centering
  \includegraphics[width=.98\linewidth]{precfom_py.pdf}
  \caption{Error estimates of the approximation as defined in~\eqref{eq:deffom} for the precession modulations, for the three cases $++$, $\perp\perp$ and $--$ listed in Table~\ref{tab:precexamples}. The thick and thin vertical lines show, respectively, the merger frequency and the asymptotic ringdown frequency. The colors correspond to different values of $m$ in $\calT^{2}_{m 2}$ as defined in~\eqref{eq:defprectransfer}. The full lines are computed with numerical derivatives for the full waveform and modulation, while in the cases $++$ and $\perp\perp$ the dashed lines show the leading-order PN estimates~\eqref{eq:precfomPN} for single-spin simple precession. We show the three error estimates $\epsilon_{\Psi 2}$, $\epsilon_{A1}$, $\epsilon_{A2}$ defined in~\eqref{eq:deffom}, and the range $\epsilon \gtrsim 1$ indicates a breakdown of the formalism of Sec.~\ref{sec:formalism}.}
  \label{fig:fomprec}
\end{figure*}

@@

\begin{figure*}
  \centering
  \includegraphics[width=.98\linewidth]{trigopoly_py.pdf}
  \caption{Amplitude (left panel) and phase (right panel), compared to its trigonometric polynomial representation for the three cases listed in Table~\ref{tab:precexamples}. The thick and thin vertical lines indicate the frequency of the merger and ringdown (the QNM frequency) respectively. The continuous line shows the target signal $\tilde{h}(f)$, while the dashed line shows the artificially symmetrized trigonometric polynomial~\eqref{eq:hsymtrigo}. The dots show the discrete samples entering~\eqref{eq:ykDFTprec}. The lower panels show the amplitude and phase residuals of~\eqref{eq:hsymtrigo} compared to the original $\tilde{h}(f)$.}
  \label{fig:trigopoly}
\end{figure*}

@@

\begin{figure*}
  \centering
  \includegraphics[width=.98\linewidth]{precerror_++_py.pdf}
  \caption{Fourier-domain amplitudes of the modes $\tilde{h}^{\rm I}_{\ell m}$ (top panels), reconstruction errors in the transfer fonctions $\calT^{2}_{m2}$ for the amplitudes (middle panels) and for the phases (bottom panels),  for the configuration $++$ of Table~\ref{tab:precexamples}, with spins almost aligned. The thick and thin vertical lines show the merger and asymptotic ringdown frequencies respectively. For the solid curves in the top panel, transfer functions were computed from an FFT of the full time-domain precessing signal, while the dashed curves show the results of the Fourier-domain reconstruction techniques presented here. The columns show the orders of approximation $\{N:0 | A:0 | \text{No Conv.}\}$, $\{N:3 | A:2 | \text{No Conv.}\}$, $\{N:3 | A:2 | \text{Conv.}\}$. In the right column, the region of transition between the perturbative treatment of Sec.~\ref{subsec:executivesummary} and the high-frequency treatment of Sec.~\ref{subsec:trigopoly} is shaded in grey. Note that the errors shown here are relative to a given mode, not to the leading mode, and the errors are largest in this sense for the most subdominant modes; the top panels illustrate the mode hierarchy.}
  \label{fig:precerrors++}
\end{figure*}

\begin{figure*}
  \centering
  \includegraphics[width=.98\linewidth]{precerror_pp_py.pdf}
  \caption{Same as Fig.~\ref{fig:precerrors++}, but for the configuration $\perp\perp$ of Table~\ref{tab:precexamples}, with both spins in the orbital plane.}
  \label{fig:precerrorspp}
\end{figure*}

\begin{figure*}
  \centering
  \includegraphics[width=.98\linewidth]{precerror_--_py.pdf}
  \caption{Same as Fig.~\ref{fig:precerrors++}, but for the configuration $--$ of Table~\ref{tab:precexamples}, with almost anti-aligned spins.}
  \label{fig:precerrors--}
\end{figure*}

\begin{figure*}
  \centering
  \includegraphics[width=.98\linewidth]{precMM_py.pdf}
  \caption{Unfaithfulness at various orders of approximation as compared with an FFT of the full time-domain precessing signal, for the three precessing systems of Table~\ref{tab:precexamples}. The color indicates the order of approximation of the perturbative treatment of Sec.~\ref{subsec:executivesummary}. Dashed lines are obtained applying the treatment of Sec.~\ref{subsec:executivesummary} to the whole frequency band, while full lines are obtained using Sec.~\ref{subsec:trigopoly} for the high frequency range. The computation is done using~\eqref{eq:defMM} with $f_{\rm min} = 20\mathrm{Hz}$, for masses between 20 and 400 $\Msol$, and with the aLIGO ZDHP noise curve~\cite{LIGOProspects13}. The inclination is defined here as the angle between the line of sight and the final $\bm{J}$.}
  \label{fig:precunfaithfulness}
\end{figure*}
